\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}

\newcommand{\code}[1]{\texttt{#1}}

\begin{document}

\title{Practical assignment 0: Environment}
\author{Anatolii Stehnii\\Introduction to databases}
\maketitle

This assignment is an instruction on the environment setup for the following practical assignments. I recommend you to do this assignment before the start of practical lessons.

\begin{enumerate}
\item Install IDE \href{https://www.jetbrains.com/rider/download/}{JetBrains Rider}. If you have not done this yet, obtain a JetBrains student license \href{https://www.jetbrains.com/student/}{here} and activate the Rider.
\begin{enumerate}
\item Install plugin \href{https://plugins.jetbrains.com/plugin/7017-plantuml-integration}{PlantUML} to render UML diagrams.
\item Install \href{http://www.graphviz.org/download/}{Graphviz}.
\end{enumerate}
\item Install \href{https://www.microsoft.com/net/learn/get-started}{.NET Core 2.0}. Instructions for different Linux repositories can be found  \href{https://docs.microsoft.com/ru-ru/dotnet/core/linux-prerequisites?tabs=netcore2x}{here}.
\begin{enumerate}
\item Install \href{https://github.com/dotnet/core/blob/master/release-notes/download-archives/1.0.4-sdk-download.md}{.NET Core SDK}.
\end{enumerate}
\item Install \href{https://www.mysql.com/}{MySQL} (server 5.7 community edition; instructions for \href{https://www.digitalocean.com/community/tutorials/how-to-install-mysql-on-ubuntu-16-04}{Ubuntu}). Kindly advise you to remember the root password.
\begin{enumerate}
\item For Windows, you need to download the \href{https://dev.mysql.com/downloads/installer/}{installer} (the second option), run the .exe file, for products choose only MySQL Server 5.7.21, and then for installation only choose the MySQL Workbench 6.3.10. Do not forget to add the path to an executable \code{mysql.exe} to \code{PATH} variable.
\end{enumerate}
\item Fork and clone git repository \href{https://github.com/tsdaemon/ucubot}{ucubot} (you have to clone your version of the repository).

\end{enumerate}

If you have any issues with this assignment, please reach me in Slack @Anatolii Stehnii or send email to \href{mailto:stehnii@ucu.edu.ua}{stehnii@ucu.edu.ua}. Feel free to use course Slack channel to ask questions and help your colleagues. Environment setup always raises questions, so do not hesitate to ask for my help. If you have any new ideas or improvements, feel free to send pull-request to repository \href{https://github.com/tsdaemon/ucu-databases-course}{ucu-databases-course}. Your colleagues will be grateful if you make the instructions better.

\end{document}
